---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# SpiGesXML

<!-- badges: start -->
[![swissstatsr badge](https://swissstatsr.r-universe.dev/badges/:name)](https://swissstatsr.r-universe.dev/)
[![SpiGesXML status badge](https://swissstatsr.r-universe.dev/badges/SpiGesXML)](https://swissstatsr.r-universe.dev/SpiGesXML)
<!-- badges: end -->

The goal of SpiGesXML is to ease extraction of data from SpiGes XML files.

## Installation

``` r
# install from the r-universe
install.packages("SpiGesXML", repos = "https://swissstatsr.r-universe.dev")
```

``` r
# install from GitHub
pak::pak("SwissStatsR/SpiGesXML")
```

## Examples

```{r}
library(SpiGesXML)

# SpiGes XML file example 
# https://www.bfs.admin.ch/bfs/de/home/statistiken/gesundheit/gesundheitswesen/projekt-spiges.assetdetail.27905035.html
xml_example <- "https://dam-api.bfs.admin.ch/hub/api/dam/assets/27905035/master"

spiges_get_df(x = xml_example, node = "Administratives")
```

Variables can be individually selected using the `variables` argument:

```{r}
spiges_get_df(x = xml_example, node = "Administratives", variables = c("abc_fall", "geschlecht"))
```


Show node names available for data extraction in XML SpiGes:

```{r}
spiges_get_name_nodes(x = xml_example)
```

Get list of variable names available by SpiGes node names:

```{r}
spiges_get_name_variables() |>
  str()
```

## Notes

The following node names are not available for data extraction: "Unternehmen", "Standort", "Fall" and "Kantonsdaten".

## TODO

- Read-in the "Kantonsdaten" (on levels "Fall", "Standort", "Unternehmen").
    - There are no specified attributes as each canton can set its own variables... Could be problematic for including it in the function.
- Use the schema-File (".xsd") to identify the variables instead of hard-coding it. 

